---
title: "Project 8"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


Data

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source:http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. 

Goal of project

Predict the manner in which they did the exercise.

Library and Data load
```{r}
## Loading required libraries
library(caret)
library(randomForest)

#Downloading data
if (!file.exists('train.csv')) {
  download.file(url = 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', 
                destfile = 'train.csv', method = 'curl', quiet = TRUE) 
}
if (!file.exists('test.csv')) {
  download.file(url = 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv', 
                destfile = 'test.csv', method = 'curl', quiet = TRUE)
}
trainRaw <- read.csv('train.csv')
testRaw <- read.csv('test.csv')

```

Data Preprocessing
```{r}
#removing unrelated columns such as column number and time stamp
str(trainRaw)
train <- trainRaw[, 6:ncol(trainRaw)]

#Train and test data set creation
set.seed(23954)

inTrain <- createDataPartition(y = train$classe, p = 0.7, list = F)
training <- train[inTrain, ]
testing<- train[-inTrain, ]

#removing simillar variables
nzv <- nearZeroVar(train, saveMetrics = T)
keepFeat <- row.names(nzv[nzv$nzv == FALSE, ])
training <- training[, keepFeat]

#removing variables with all NAs
training <- training[, colSums(is.na(training)) == 0]
dim(training)


#removing simillar variables
nzv <- nearZeroVar(testing, saveMetrics = T)
keepFeat <- row.names(nzv[nzv$nzv == FALSE, ])
testing <- testing[, keepFeat]

#removing variables with all NAs
testing <- testing[, colSums(is.na(testing)) == 0]
dim(testing)

```

Modeling
```{r}
# 5 fold cross validation
modCtl <- trainControl(method = 'cv', number = 5,  verboseIter = TRUE, allowParallel = TRUE)

#random forest modeling
set.seed(2384)
modRf <- train(classe ~. , data = training, method = 'rf', trControl = modCtl, verbose = TRUE)
modRf$finalModel
```

```{r}
#Test data error check
predRf <- predict(modRf, newdata = testing)
confusionMatrix(predRf, as.factor(testing$classe))$table
confusionMatrix(predRf, as.factor(testing$classe))$overall[1]
```

We got 99% accuracy with the Random forest model with 5 fold cross validation

```{r}
#Predicting on Test dataset
predRfTest <- predict(modRf, newdata = testRaw)
predRfTest
```

