---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
library(data.table)
library(ggplot2)

data <- data.table::fread(input = "data/activity.csv")
```


```{r}
head(data)
```

## What is mean total number of steps taken per day?

```{r}
Total_Steps_day <- data[, c(lapply(.SD, sum, na.rm = FALSE)), .SDcols = c("steps"), by = .(date)] 

ggplot(Total_Steps_day, aes(x=steps)) + 
  geom_histogram(binwidth = 1000)+
  labs(title = "Daily Steps", x = "Steps", y = "Frequency")

```

```{r}
paste0("Mean: ", mean(Total_Steps_day$steps, na.rm = TRUE))
paste0("Median: ", median(Total_Steps_day$steps, na.rm = TRUE))
```

## What is the average daily activity pattern?
```{r}
mean_daily_steps <- data[, c(lapply(.SD, mean, na.rm= TRUE)), .SDcols = c("steps"), by = .(interval)]

ggplot(mean_daily_steps, aes(x= interval, y = steps)) +
  geom_line(type = "1") +
  labs(title = "Average Daily Steps", x = "Interval", y= "Mean")

```


```{r}
mean_daily_steps[steps == max(steps), "interval"]
```
## Imputing missing values

1. Finding total number of missing values in data 
```{r}
data[is.na(steps), .N]
```

2. To impute the missing values we can fill it with median values

```{r}
filled_data <- data
filled_data[is.na(steps), "steps"] <- filled_data[, c(lapply(.SD, median, na.rm = TRUE)),  .SDcols = c("steps")]
filled_data[is.na(steps), .N]
```

```{r}
total_steps_day <- filled_data[, c(lapply(.SD, sum)), .SDcols = c("steps"), by = .(date)]
```

new daily mean and median

```{r}
paste0("Mean: ", mean(total_steps_day$steps, na.rm = TRUE))
paste0("Median: ", median(total_steps_day$steps, na.rm = TRUE))
```

```{r}
ggplot(total_steps_day, aes(x=steps)) + 
  geom_histogram(binwidth = 1000)+
  labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
filled_data[, date := as.POSIXct(date, format = "%Y-%m-%d")]
filled_data[, `Day of Week`:= weekdays(x = date)]
filled_data[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", x = `Day of Week`), "weekday or weekend"] <- "weekday"
filled_data[grepl(pattern = "Saturday|Sunday", x = `Day of Week`), "weekday or weekend"] <- "weekend"
filled_data[, `weekday or weekend` := as.factor(`weekday or weekend`)]
head(filled_data, 10)
```

```{r}
filled_data[is.na(steps), "steps"] <- filled_data[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
IntervalDT <- filled_data[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `weekday or weekend`)] 

ggplot(IntervalDT , aes(x = interval , y = steps, color=`weekday or weekend`)) + geom_line() + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2)
```



  