---
title: "Project 2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data
```{r}
data(ToothGrowth)
```

## Analysis
### Setup the environment
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(datasets)
data(ToothGrowth)
```

### Exploratory data analysis

```{r}
ToothGrowth %>% group_by(supp, dose) %>%
  summarise(meanLength = mean(len), medianLength = median(len), sdLength = sd(len))
```

Observations: 

 - Guinea pigs treated with 0.5 and 1mg/day of orange juice have on average longer tooth length than those treated with vitamin C. However, this trend is not obvious in guinea pigs treated with 2mg/day of orange juice vs vitamin C
 
2. Plot the distribution of the raw data  
```{r, echo=FALSE}
ggplot(data = ToothGrowth, aes(x = supp, y = len, fill = supp)) +
  geom_boxplot() +
  geom_point() +
  facet_grid(. ~ dose) +
  xlab('Supplement type') +
  ylab('Tooth length') +
  ggtitle('Tooth growth of varying doses and supplement type in guinea pigs') +
  scale_fill_brewer(palette = 'Pastel1',
                    labels = c('Orange juice', 'Vitamin C'),
                    name = 'Supplement type')
```


### Hypothesis testing: difference in tooth growth treated with orange juice and vitamin C
- Null hypothesis: there is no difference in tooth growth in guinea pigs treated with the same dose of orange juice and vitamin C
- Alternative hypothesis: there is a difference in tooth growth in guinea pigs treated with the same dose of orange juice and vitamin C
- Control type I error at less than 0.05 (p<0.05)

1. Stratify data by three doses: low dose (0.5mg/day), medium dose (1mg/day), high dose (2mg/day)
```{r}
ld <- ToothGrowth[ToothGrowth$dose == 0.5, ]
md <- ToothGrowth[ToothGrowth$dose == 1, ]
hd <- ToothGrowth[ToothGrowth$dose == 2, ]
```

2. Perform hypothesis testing for low dose (0.5mg/day)
```{r}
t.test(ld$len ~ ld$supp, alternative = 'two.sided', paired = F, var.equal = F)$p.value
```
  As p-value is less thean type I error(0.05), we can reject null hypothesis.
  

3. Perform hypothesis testing for medium dose (1mg/day) - code not shown but similar to above
```{r, echo = F}
t.test(md$len ~ md$supp, 
       alternative = 'two.sided',
       paired = F, var.equal = F, conf.level = 0.95)$p.value
```
  As p-value is less thean type I error(0.05), we can reject null hypothesis.

4. Perform hypothesis testing for high dose (2mg/day) - code not shown
```{r, echo = F}
t.test(hd$len ~ hd$supp, 
       alternative = 'two.sided',
       paired = F, var.equal = F, conf.level = 0.95)$p.value
```
  The mean tooth length when treated at 2mg/day with orange juice is not significantly different than that of vitamin C.

### Hypothesis testing: difference in tooth growth treated with different doses of supplements
- Null hypothesis: there is no difference in tooth growth in guinea pigs treated with different doses of orange juice or vitamin C
- Alternative hypothesis: there is a difference in tooth growth in guinea pigs treated with different doses of orange juice or vitamin C
- Control the type I error rate at less than 0.05 (p<0.05)

1. Stratify data by two supplement types, orange juice and vitamin C, and then strtify the data to contain only two doses (since t.test can only take conlumns containing two levels) - code not shown
```{r, echo = F}
oj <- ToothGrowth[ToothGrowth$supp == 'OJ', ]
oj_lm <- oj[oj$dose <= 1, ]
oj_lh <- oj[oj$dose != 1, ]
oj_mh <- oj[oj$dose >= 1, ]
vc <- ToothGrowth[ToothGrowth$supp == 'VC', ]
vc_lm <- vc[vc$dose <= 1, ]
vc_lh <- vc[vc$dose != 1, ]
vc_mh <- vc[vc$dose >= 1, ]
```
2. Perform hypothesis testing for orange juice at different doses
Three different doses are compared to each other. The family-wise error rate will be controlled using the Bonferroni correction to adjust the p-value
```{r}
oj_lm_p <- t.test(oj_lm$len ~ oj_lm$dose, alternative = 'two.sided', paired = F, var.equal = F)$p.value
oj_lh_p <- t.test(oj_lh$len ~ oj_lh$dose, alternative = 'two.sided', paired = F, var.equal = F)$p.value
oj_mh_p <- t.test(oj_mh$len ~ oj_mh$dose, alternative = 'two.sided', paired = F, var.equal = F)$p.value
data.frame(comparisons = c('0.5 vs 1', '0.5 vs 2', '1 vs 2'),
           pvalue = c(oj_lm_p, oj_lh_p, oj_mh_p),
           bonferroni = p.adjust(c(oj_lm_p, oj_lh_p, oj_mh_p), method = 'bonferroni'))
```
  
  
  Since the Bonferroni-corrected p-values for 0.5 vs 1 and 0.5 vs 2 comparisons are less than 0.05, the null hypotheses are rejected. 

3. Perform hypothesis testing for vitamin C at different doses

Three different doses are compared to each other. The family-wise error rate will be controlled using the Bonferroni correction to adjust the p-value - code not shown
```{r, echo = F}
#low vs medium dose
vc_lm_p <- t.test(vc_lm$len ~ vc_lm$dose, alternative = 'two.sided', 
                  paired = F, var.equal = F, conf.level = 0.95)$p.value
#low vs high dose
vc_lh_p <- t.test(vc_lh$len ~ vc_lh$dose, alternative = 'two.sided',
                  paired = F, var.equal = F, conf.level = 0.95)$p.value
#medium vs high dose
vc_mh_p <- t.test(vc_mh$len ~ vc_mh$dose, alternative = 'two.sided',
                  paired = F, var.equal = F, conf.level = 0.95)$p.value
data.frame(comparisons = c('0.5 vs 1', '0.5 vs 2', '1 vs 2'),
           pvalue = c(vc_lm_p, vc_lh_p, vc_mh_p),
           bonferroni = p.adjust(c(vc_lm_p, vc_lh_p, vc_mh_p), method = 'bonferroni'))
```
  Since the Bonferroni-corrected p-values for all comparisons are less than 0.05, the null hypotheses are rejected. 

### Conclusions
1. The assumptions made to perform the statistical tests

  - At different dose level there is a different variance in Orange juice and Vitamin-C treated groups
  
  - Variance are different in the orange juice- and vitamin C-treated groups at different dose levels
  - The different treatment groups are non-paired - this is not so much an assumption but how the experiment is actually condudcted, as specified in the **?ToothGrowh** description
  - Assume all the observations are independent and identically distributed (iid)
  - Assume tooth growht follows normal distribution

2. Conculsions
  - At doses of 0.5mg/day and 1mg/day, guinea pigs treated with orange juice have longer tooth length compared to those treated with vitamin C; however, the difference is not significant at 2mg/day
  - In guinea pigs treated with orange juice, treatment at 1mg/day and 2mg/day results in on average longer tooth length than those treated at 0.5mg/day; however, the difference is not significant in 1mg/day compared to 2mg/day treatment groups
  - In guinea pigs treated with vitamin C, there is a does-dependent effect in that the higher the dose is, the longer the tooth length is. Whether this relationship is linear requires further exploration.
